---
title: "1008 class"
author: "River Sung"
date: "08/10/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data types and structures
## Core data types
Aka "modes"  
`typeof` - determine the mode of the object  
`mode` - also determine the mode of the object, but cannot distinguish between **integers** and **doubles**

### Numeric
**Numeric** data type is the simplest. It contains **integers** or **doubles**  
**Integers** - whole numbers  
**Doubles** - decimal numbers 
```{r}
x <- c(1.0, -3.4, 2, 140.1) # numeric and double

typeof(x) #using this function to determine what mode/data type x is 
mode(x) #does not distinguish integers and doubles 

#Removing decimal does not make a double into an integer. 
#For example: 
typeof(5)
```
Therefore, for R to recognize a value as integer, we add a capital L
```{r}
x <- 5L
typeof(x)
```

### Character 
Mostly seen as letters or words. 
```{r}
x <- c("bubul", "magpie", "spoonbill", "barbet")
typeof(x)

# Can also be two numeric vector 
x <- 3
y <- 5.3
x + y
```
When " " is used, R interprets them as characters, but some functions like  `plot(x, y)` automatically converts characters to doubles. 
```{r}
x <- "3"
y <- "5.3"

# x+ y
# Error in x + y : non-numeric argument to binary operator
plot(x, y)
```

### Logical 
`TRUE` or `FALSE` conditions, also `1` or `0`
```{r}
# full text:
x <- c(TRUE, FALSE, FALSE, TRUE)
# or in short:
x <- c(T, F, F, T)
# or binary 
x1<-c(1,0,0,1)
x2 <- as.logical(c(1,0,0,1)) #converts internally to T or F
# OR: x3 <- as.logical(c(1,0,0,1))
```

## Derived data types 
Stored as numeric, character, or logical, but have additional attributes allowing to be handled certain way. These attributes define the `class` of an object. 
### Factor
Used to group variables into fixed number of unique categories or **levels**. Most time data type must be converted into a factor using `as.factor` 
```{r}
a <- c("M", "F", "F", "U", "F", "M", "M", "M", "F", "U")
typeof(a) # mode character
class(a) # class character

a.fact <- as.factor(a)
class(a.fact) # class factor
```
**Note**: a has not changed the mode. The data is still of the type `character`. However, it may seem strange at first, but the derived object `a.fact` is now stored as an `integer`!
```{r}
mode(a.fact)
typeof(a.fact)
a.fact
```
`a.fact` is stored as `interger`, even though it's displayed as factors (characters), because `a.fact` is more complicated. `Factor`stores additional information not visible in output. Hidden information is stored in `attributes`
```{r}
attributes(a.fact)
levels(a.fact)
```
There are two attributes: `levels` and `class`. The `levels` attribute lists unique values in alpha-numeric order. **If we need to rearrange for display and graphing purposes, must change it by recreating the factor object:**
```{r}
factor(a, levels=c("U", "F", "M"))
```

Practice 3.1
```{r}
str(iris)
iris.sel <- subset(iris, Species == "setosa" | Species == "virginica") #subsetting thw two species 
iris.sel
levels(iris.sel$Species) # even though it's been subset, the third species from original dataset is still present
boxplot(Sepal.Width~Species, data=iris.sel) 
```
  
Plotting graph allows us to visualize that there's an empty column, which is bad. Need to get rid of it by using the `droplevels` function.
```{r}
iris.sel$Species <- droplevels(iris.sel$Species) 
iris.sel
levels(iris.sel$Species) # now that there are only two levels present in the dataset 
boxplot(Sepal.Width~Species, data=iris.sel) # graph fixed! 
rownames(iris.sel) = seq(length=nrow(iris.sel)) #wait i'm not sure what this is for
```

### Date 
Values stored as numbers. To be properly interpreted as a date object in R, their attribute must be explicitly defined as **date**. There is a package called `lubricate`. This topic will not be covered.  

### NAs and NULLs
Missing or unknown values in dataset does not equal to 0.  
NA - not available: missing value, data should be there, but we are unable to retrieve it.  
NULL - also missing values, but means value does not exist or not measurable.   
When doing analysis, this could lead to very different results. 
```{r}
x <- c(23, NA, 1.2, 5)
y <- c(23, NULL, 1.2, 5)
mean(x) # cannot take the mean of NA values
mean(y)
```
Can use the `na.rm` function to ignore NA values 
```{r}
mean(x, na.rm = T) #works! now can calculate the mean even if the value is NA 
```

## Data Structures 
There are different such as **table** or a **list** of results. They are stored in R in one of several data structures. 

### (Atomic) vectors
Consist of an ordered set of values of the same type or class (numeric, character, date). **Vector can be created with the combination function `c()`**
```{r}
x <- c(674 , 4186 , 5308 , 5083 , 6140 , 6381)
x
x[3] # access the third element of x, use [ ]
x[c(1,3,4)] # select a subset of elements by index values using `c()`
x[2:4] # select a range of element like index 2-4 use the operator :
x[2] <- 0 # assign value to the second index - replace 
x
x <- c("all", "b", "olive") # can store other data type, like characters 
x
x <- c( 1.2, 5, "Rt", "2000") # can only be of one type, cannot mix numeric and character type 
typeof(x)
```
When more than two data types are given at once, highest common mode is used. **NULL < logical < integer < double < character**  

### Matrices and arrays
Vectors that are indexted with two indices instead of one. Following: 3x3 matrix with generated values. `nrow` and `ncol` define dimension and `runif` generates 9 values between 0-10
```{r}
m <- matrix(runif(9,0,10), nrow = 3, ncol = 3)
m
```
Higher dimensions, use `array` function to create n-dimensional matrix. 3x3x3 can be created. 
```{r}
m <- array(runif(27,0,10), c(3,3,3))
m
```
Matrices and arrays can store numeric or character data types, but not both at once. 

### Data frames 
Close to a table, but can mix data types across columns (both numeric and character columns can coexist in a data frame). Data type same across rows 
```{r}
name   <- c("a1", "a2", "b3")
value1 <- c(23, 4, 12)
value2 <- c(1, 45, 5)
dat    <- data.frame(name, value1, value2)
dat
str(dat)
attributes(dat)
names(dat) # extract column names 
rownames(dat) # extract row names 
```

### Lists
An ordered set of components stored in 1D vector. It's a **recursive vector**, where each vector element can have a different data type and structure. Each element of a list can contain complex objects such as matrix, data frames, or another list.  
  
List can be created with the `list` function. Example:  
 - A: Two-column data frame  
 - B: Logical vector with two elements   
 - C: Character vector with three elements   
```{r}
A <- data.frame(
     x = c(7.3, 29.4, 29.4, 2.9, 12.3, 7.5, 36.0, 4.8, 18.8, 4.2),
     y = c(5.2, 26.6, 31.2, 2.2, 13.8, 7.8, 35.2, 8.6, 20.3, 1.1) )
B <- c(TRUE, FALSE)
C <- c("apples", "oranges", "round")
my.list <- list(A = A, B = B, C = C)
my.list

names(my.list)
```
Each component of a list can be extracted using the $ symbol followed by that component’s name. For example, to access component A from list `my.list`, type:  
```{r}
my.list$A

# or using its numerical index 

my.list[[1]] # first (1) component of the list 

# can also create list like this: 
my.list.notags <- list(A, B, D)
my.list.notags

names(my.list.notags) # do not have component names NULL
```
If we fit a linear model between elements x and y (in the data frame A), the output of this model (M) is a list:
```{r}
M <- lm( y ~ x, A)
str(M)
names(M)
str(M$qr) # the element qr is a list with 5 elements 
M$qr$rank # access component qr of the list M
```

## Coercing data 
Data can be converted from one type to another. From character -> numeric, use `as.numeric`.  
This function converts the vector into a double (can also use `as.double`). If convert y into an integer, R remove all decimals from the number. 
```{r}
y <- c("23.8", "6", "100.01","6")
y.c <- as.numeric(y)
y.c
as.integer(y) # removes decimals when converting to integer 

numchar <- as.character(y.c) # convert a number to a character 
numchar

numfac <- as.factor(y) # convert a number/character into factor 
numfac

charfac <- as.factor(y.c)
charfac
```

Other functions: 
```{r}
as.numeric  # Coerce to numeric
as.double   # Coerce to double
as.integer  # Coerce to integer
as.character #  Coerce to character
as.logical  # Coerce to Boolean (logical: TRUE | FALSE)
as.factor   # Coerce to factor
as.Date  # Coerce to date
as.data.frame  # Coerce to data frame
as.list # Coerce to list
```



