---
title: "1008 class"
author: "River Sung"
date: "08/10/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data types and structures
## Core data types
Aka "modes"  
`typeof` - determine the mode of the object  
`mode` - also determine the mode of the object, but cannot distinguish between **integers** and **doubles**

### Numeric
**Numeric** data type is the simplest. It contains **integers** or **doubles**  
**Integers** - whole numbers  
**Doubles** - decimal numbers 
```{r}
x <- c(1.0, -3.4, 2, 140.1) # numeric and double

typeof(x) #using this function to determine what mode/data type x is 
mode(x) #does not distinguish integers and doubles 

#Removing decimal does not make a double into an integer. 
#For example: 
typeof(5)
```
Therefore, for R to recognize a value as integer, we add a capital L
```{r}
x <- 5L
typeof(x)
```

### Character 
Mostly seen as letters or words. 
```{r}
x <- c("bubul", "magpie", "spoonbill", "barbet")
typeof(x)

# Can also be two numeric vector 
x <- 3
y <- 5.3
x + y
```
When " " is used, R interprets them as characters, but some functions like  `plot(x, y)` automatically converts characters to doubles. 
```{r}
x <- "3"
y <- "5.3"

# x+ y
# Error in x + y : non-numeric argument to binary operator
plot(x, y)
```

### Logical 
`TRUE` or `FALSE` conditions, also `1` or `0`
```{r}
# full text:
x <- c(TRUE, FALSE, FALSE, TRUE)
# or in short:
x <- c(T, F, F, T)
# or binary 
x1<-c(1,0,0,1)
x2 <- as.logical(c(1,0,0,1)) #converts internally to T or F
# OR: x3 <- as.logical(c(1,0,0,1))
```

## Derived data types 
Stored as numeric, character, or logical, but have additional attributes allowing to be handled certain way. These attributes define the `class` of an object. 
### Factor
Used to group variables into fixed number of unique categories or **levels**. Most time data type must be converted into a factor using `as.factor` 
```{r}
a <- c("M", "F", "F", "U", "F", "M", "M", "M", "F", "U")
typeof(a) # mode character
class(a) # class character

a.fact <- as.factor(a)
class(a.fact) # class factor
```
**Note**: a has not changed the mode. The data is still of the type `character`. However, it may seem strange at first, but the derived object `a.fact` is now stored as an `integer`!
```{r}
mode(a.fact)
typeof(a.fact)
a.fact
```








```{r}
iris
str(iris)

iris.sel <- subset(iris, Species == "setosa" | Species == "virginica")
iris.sel
levels(iris.sel$Species)

iris.sel$Species <- droplevels(iris.sel$Species)
iris.sel

```
